Cover letter for the original IJRR submission

### 2017-10-06 reviews_tro/reviewer_response.tex

\quot{* I am not sure why the $C_{reach}$ hypothesis is described as the ``strong''
hypothesis in, e.g., Sect. IV. Is there a corresponding weak
hypothesis?}
\as{We apologize for the confusion, that comes from an inaccurate use of English. What we meant by ``strong'' is that the hypothesis is constraining, because
it defines a solution space different from the real one, and that we need to demonstrate that it makes sense.  We removed ``strong'' and replaced it by ``main'' to clarify. }\done

@@

\quot{* In Sect. V.A., the authors mention that the ability to plan a contact
break and creation with the same limb allows the creation of dynamic
motions. This is quite an important ability for an otherwise
quasi-static planner, but this ability isn't described again until the
appendix. It would be worth explaining why this is true in the paper
itself. }
\as{Thanks for the suggestion. We added a paragraph to better explain this in Section V.A.}
\qt{This representation is sufficient to describe all the contact phases, because the single support phase is implicitely described. Furthermore it removes the need to compute a single-support quasi-static configuration as in the example. Indeed, there might be a case where no quasi-static solution exists for the single support phase (because of the environment), but
there exists a dynamic motion connecting the two double support states. Such motion will be computed by our framework, because the quasi-static constraint is only required at the contact planning phase; as shown in the companion video, and explained in Appendix C, our framework is able to produce dynamic motions.
}
\done

\quot{In Sect. V.B, the authors mention inserting an additional state
between contact configurations if two or more contacts cannot be
maintained. Shouldn't this step appear somewhere in Alg. 1?}
\as{We apologize for the lack of clarity. This is actually handled by the method \textsc{RepositionContacts}. We now realize that the name is misleading. Indeed,
\textsc{RepositionContacts} either adds a new contact, or tries to reposition one already existing. In Appendix D, we state:
\textit{Otherwise, the method \textsc{RepositionContacts} is called. It repositions one end effector (either a free
limb, or the oldest active contact) towards a new contact
position if possible. }

We changed the name to \textsc{IntermediateContactState}, with the hope that it makes things clearer.}\done

\quot{* In Appendix A.A, it is mentioned that step 2 in the workspace
computation is conservative, with reference to figure 17-right. But
Fig. 17-right is the simplification of the convex hull, corresponding
to step 3, and it's actually hard to tell from the figure that it is
conservative. Not knowing anything about Blender's decimate tool, I
assume it is operating by removing vertices, which would indeed be
conservative, but some clarification would be useful here.}
\as{Indeed, the decimate tool removes vertices, and this was not explained clearly in the paper. We updated the text accordingly.}
\qt{The resulting polytope can contain a very large number of faces. A last step is thus to simplify it with the blender decimate tool (\url{http://wiki.blender.org/index.php/Doc:2.4/Manual/Modifiers/Generate/Decimate}). This tool removes a user-defined amount of vertices (and faces) of the polytope, thus resulting
in a convervative approximation of the original shape. For HRP-2 we apply the operator with a ratio of $0.06$, resulting in a polytope of 38 faces for the arms and the legs.}\done

### 2017-10-06 reviews_tro/reviewer_response.tex

\documentclass[a4paper]{article}
\usepackage{fullpage}
\usepackage{amsfonts} % pour les lettres maths creuses \mathbb
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[frenchb]{babel}
\usepackage{aeguill}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{color}
\usepackage{listings}
\usepackage{pdfpages}
% Another method to track changes
% Examples of usage:
% - This is \added[id=per,remark={we need this}]{new} text.
% - This is \deleted[id=per,remark=obsolete]{unnecessary}text.
% - This is \replaced[id=per]{nice}{bad} text.
% To print the list of change use: \listofchanges
%~ \usepackage{changes}	% use this for the working version
%\usepackage[final]{changes} % use this for the final version
%~ \definechangesauthor[name={Andrea Del Prete}, color=orange]{adp}
\newcommand{\deladp}[1]{\textcolor{red}{#1}}
\newcommand{\addadp}[1]{\textcolor{green}{#1}}
%~ \newcommand{\repadp}[2]{\replaced[id=adp]{#1}{#2}}
%~ \definechangesauthor[name={Steve Tonneau}, color=blue]{st}
%~ \newcommand{\delst}[1]{\deleted[id=st]{#1}}
%~ \newcommand{\addst}[1]{\added[id=st]{#1}}
%~ \newcommand{\repst}[2]{\replaced[id=st]{#1}{#2}}
\newcommand{\gls}[1]{\textit{#1}}
\newcommand{\glslink}[2]{{#2}}
%~ \newcommand{\done}[0]{\textcolor{green}{DONE}}
\newcommand{\done}[0]{}
\newcommand{\ndone}[0]{\textcolor{red}{TODO}}

@@

\as{Agreed. We removed the incriminated text and provided an intuition of the heuristic in the text, as mentioned above.}\done

\quot{github link seems to be dead? (The link on p. 15 is OK\dots should they be the same?)}

\as{We are suprised by this comment because after testing it, the link seems to be valid. In any case, in accordance with reviewer 6, we changed all the links source codes to provide a SHA link that will bring the user to the accurate revision of the projects considered.}
\qt{(Section VII) Both HPP and our planner can be simply downloaded and compiled by following the instructions on
\url{https://humanoid-path-planner.github.io/hpp-doc/download.html?branch=rbprm}. \\
(Appendix C) To address the other scenarios, we propose a new implementation, entirely open source (\url{https://github.com/stonneau/python_sandbox/releases/tag/tro_paper} ), which can be integrated directly in our motion planner software. }\done

\quot{p.5, on right, ``less than epsilon'' between end effector and world - but doesn’t the orientation of the end effector matter tremendously here, too?}
\as{
We agree that the orientation of the end effector matters. Indeed, we could consider to further filter the set of candidates samples regarding orientation, to reduce the number of failed
projections. However, because of the joint limits constraints, computing a relevant orientation distance is less trivial than in the 3D case. We tried to implement this
in earlier versions of the planner, and did not obtain significant gains in the performance. We believe it is explained by the fact that the majority of unfeasible samples is already
efficiently rejected by the simple ``epsilon distance'' check. We updated the text of the paper to explain this.
}
\qt{Because the distance computation $\epsilon$ does not account for the variation in orientation, several samples of $ C_{Contact}^{\epsilon}$  may turn out to be unfeasible at the time of projection. One could consider additionally filtering $ C_{Contact}^{\epsilon}$ based on the orientation with respect to the obstacle normal, but in our experience we did not notice
any major improvement in the computational performances of the planner, so we do not perform this additional step. }
%~ Using an epsilon is an effective way to remove many irrelevant configurations.  and
%~ we believe that more complex solutions aiming at taking into account the orientation would end up being counter productive.
 %~ Indeed, using only 3D distance, we will end up with a few configurations
%~ that can't be projected correctly, and will end up being discarded. But the majority of irrelevant configurations would already have been rejected with
%~ the efficient 3D intersection query. Introducing additional filtering . } 
\done

@@

\quot{in my opinion a T-RO regular paper should be more self-containing. In several points the authors heavily rely on previous publications, e.g. when presenting the RRT-based planner (Section IV-C). This part should be expanded, describing in more detail what exactly has been changed with respect to the reference paper. }
\as{Thanks for the comment. As it is also suggested by reviewer 1 and the AE, we modified the paper to make it self contained. The references to previous work now solely indicate
that a certain method is not a contribution of the present paper. \\ 
In the particular case of Section IV-C, in fact the code of our implementation of the Bi-RRT with respect to the original one is exactly the same, with a single difference: when
traditionally a collision detection predicate is called to validate a configuration, we replace this predicate with a call to the reachability condition. We modified the text to explain this more clearly. As the Bi-RRT is a famous standard algorithm in the literature, we do not think it is necessary to explain it in this paper, but we will gladly do so if required. }
\qt{(Section IV-C) $C_{Reach}^0$ can be sampled efficiently thanks to Eq.~5, and can thus be used with any standard motion planner.
Our current implementation uses the Bi-RRT planner [25] provided by the HPP software [26].
Our implementation is exactly the same as the pseudo-code of the original planner (which does not detail the configuration validation method). With respect to a ``classic'' implementation, the only difference is that instead of validating a configuration using collision detection, we validate it with the \textit{reachability condition}.} \done

\quot{I am a bit skeptical about the readability of the figures. Some of them seem to be too small (e.g. Fig. 3, Fig. 4), some are of quite poor quality (why the robot silhouettes in Fig. 4 are blurred ?), and the purpose and meaning of Fig. 6 is rather unclear. \\ In general, the simulation results presented qualitatively in the paper are a bit hard to interpret without the accompanying video, particularly if someone looks at a b/w print of the paper (as it may be looking in T-RO).}
\as{Thanks for the comment, with which we agree in general.  \\We increased the size of Fig. 3 and 4 as much as possible to hold on one column, because we were not satisfied with the result obtained after trying to put them on a double column. The robot silhouettes in Fig. 4 are blurred when they describe a previous state of the robots. We added this information in the text. \\ As proposed also by reviewer 1, we removed references to previous results obtained in [20], so Fig. 6 has been removed.
\\Regarding the quality of our Figures,  our simulation Figures are taken from high resolution rendering in blender. The other Figures are drawn from manually designed powerpoint vectorial graphics (converted to png). We assume that the issue encountered comes from the compression of the pdf file to fit the size constraints of the 
submission site. We tried to improve this compression, but if the paper is accepted, we believe that the uncompressed pdf obtained from the source of the paper will result in good quality pictures.
Regarding the interpretation of the results without the video, we agree with the reviewer and believe that the accompanying video is essential to the comprehension of the results. We considered three options: a) displaying several sequences of contact (as currently done in most cases); b) only displaying 2 or 3 representative consecutive contact stances (currently done
in the car scenario and in the real robot scenario, for which using a large number of pictures is not readable); c)
not providing any picture and referring to the video. We chose a mix between a) and b), because we wanted the paper to be remain as independent from the video as possible, but are willing to opt for any other option if preferred by the reviewer.}\done

@@

%~ \quot{In general, the simulation results presented qualitatively in the paper are a bit hard to interpret without the accompanying video, particularly if someone looks at a b/w print of the paper (as it may be looking in T-RO).}
%~ \as{ThAll of our simulation figures are taken from high resolution rendering in blender, so we wouldn't know how to improve their quality. If the reviewers think the paper would be better without these figures, we are willing to remove them. However, we believe that they would make the paper more dependent on the video, so we would prefer to leave them.}
%\as{\textbf{\textcolor{red}{TODO This is true, although I think he has a crappy printer because I don't think it's that bad. I don't know what to do about it though. I can't simply remove the Figures, because he tends to say he would like the paper to be more independent from the video...}}}\ndone

\quot{ I also found the description of the environment model being
used,
and the way it is used in planning quite superficial. It is said, that
``an octree data structure'' is used to keep the sampled the
end-effector positions, and it is ``intersect'' with the environment.
The idea is OK, I understand that the octree is a sample-based
representation of the effector's workspace, and using some environment
model it is possible to decide if the given configuration can provide a
contact point or not. but how is the environment described ? How
efficient is this intersection operation ? Does this method require the
environment to be described by some specific objects (e.g. polygons) ?
}

\as{Thanks for the comment. We agree that additional information must be added. In Section III we now indicate that the environment is represented as a polygon soup (or 3D mesh), and
that we assume that it is fully known.  We thus also indicate that we assume the environment and state of the robot are fully known. To perform the intersection computation, we use
the fcl collision detection library, which natively proposes the intersection method. We added a reference to the library in the paper. The performance of the method is included
in the ``collision'' column of Table II. We updated the legend of Table II to indicate that this time is negligible in comparison to the other collision detections that occur.
} 
\qt{(Section III) \textbf{The environment} $O$ is defined as the union of the obstacles $O_i$ that it contains. $O$ is represented
as a polygon soup (or mesh), where the normal of each surface is known. No further requirement is needed
by our approach. In this work, we assume the environment is fully known. State uncertainty is out of the scope of the paper. \\
(Legend of Table II)  The Collision column times includes the (negligible) octree intersection operation necessary to retrieve the candidate samples.}
\done

\quot{What about the (unavoidable) uncertainty if a world model acquired via
on-line sensing is considered ? Perhaps some of these issues (e.g.
uncertainty) are not in the scope of the paper, but I can imagine that
a robotic researcher who reads the paper and is about to use the open
source software it advertises is very interested in the answers.}
\as{That is a good observation. Although the integration of on-line sensing and state estimation in the planner is an issue currently considered by our team, it is indeed
out of the scope of this paper. As reported in the previous answer we indicated it in Section III. We think that our planner could work similarly with 
a sensing system that would output data easily interpreted as a polygon soup (as for a LIDAR), otherwise additional treatment would be required. In any case additional work would be required to plan motions robust to the uncertainty of the sensing and the estimation. } \done

\quot{I appreciate the thorough presentation of results in the paper, however
it seems that this presentation is somewhat biased towards simulation
results. Simulations are perfectly OK for planning algorithms, but if
you decided to present the experimental results (HRP-2), then at least some
illustrations should be included in the paper. now they are only in the
accompanying video.  also a comparison between the results of the same
task(s) executed in dynamic simulation and on the real robot should
convince the reader that the planning method's good performance does
not depend too much on the idealized simulation environment. }
\as{We agree with the remark, and included some illustrations of the experimental results on HRP-2 in the paper. \\
We now elaborate on the validity of the simulation with respect to an experiment on the real robot. As also suggested by reviewer 1, we added a paragraph
in the paper (Section VIII.A)  to provide more information on the simulation we use to validate our motions. This simulation is an implementation of the state of the art approaches used by the community. Of course, it does not guarantee that the motion could be executed on the robot, but it is used extensively by the community as a reasonable feasibility test. The suggestion of the reviewer to compare the simulation results with those on the real robot is interesting. However, in our simulations HRP-2 is torque controlled, whereas in our experiments HRP-2 is position controlled. This makes the usefulness of a comparison between simulation and experiments extremely doubtful, because we would be comparing motions obtained with different controllers. For these reasons, we prefer not to do this comparison.}
\qt{(Section VIII.A) To generate continuous movements from our contact plans we used either the framework proposed in [1], or our own implementation of a $\mathcal{P}_3$ solver (Appendix C).
The resulting movements have been validated either on the real HRP-2 robot (details can be found in [1]), or with our dynamic simulator, based on a state-of-the-art algorithm [25].
In the simulations we controlled the robot with a standard inverse-dynamics controller[31].
This controller tries to follow the given whole-body trajectories, giving higher priority to the center-of-mass and end-effectors tracking with respect to the joint tracking.
The controller also makes sure that the resulting contact forces lie inside the specified friction cones (we used a friction coefficient of 0.3), and that the joint position, velocity and torque limits are satisfied.
The companion video shows the obtained motions.}
\done

\section{Answers to reviewer 6}
\quot{ My primary concern, however,
is that the elements described in this paper have been described in
previous published work, which makes the particular contributions of
this work unclear. To be specific, the reachability-guided planner
appears identical to the planner in [20] (as stated by the authors in
the end of Section I.B.), and the optimal control formulation is from [1]
(again as stated in I.B.).}
\as{Thanks for the comment. Indeed, all reviewers agreed that the contribution with respect to previous work was not clearly stated. We updated the paper to reflect the following. Firstly, our planner is almost identical to the one in [20], with important additions that are the pseudo code for the algorithm (the algorithm is not given in [20]), as well as the introduction of a more efficient criterion
for checking static equilibrium (see next question). The optimal control is no longer a claimed contribution of the paper, even though it differs from [1] in its formulation. The main
novelty of the paper resides in the experimental setup and analysis, as well as the empirical proof that the planner works with real robots that are more constrained kinematically than the virtual avatars from [20].}
\qt{(Section I.B)  The present paper is an extension of our conference paper[20]. As such, the motion planning algorithm presented here is the same as the one presented in the
previous work. Section IV and V of the present paper are thus paraphrasing the original paper, with the exception of Section V.B.2, where we additionally provide the pseudo-code of the algorithm absent from the original paper. We also improve the computational performance of the method by replacing the static equilibrium test with a more efficient test based on a new LP formulation presented in Section VI.
Compared with the original paper, we also release our source code, and provide the complete procedure to use our planner for any legged robot. We thus provide
all the technical information required to use the planner.\\
To validate our approach, in this paper we had to implement or integrate other tools based on previous publications from our team. We implemented a solution to $\mathcal{P}_3$ based on a reformulation of our previous work[1]. The implementation of this solution is also provided as an open-source project, which can be integrated directly in our framework. Our solution to $\mathcal{P}_3$ then allows us to verify that the synthesized motions are physically consistent, using our implementation of a state-of-the-art simulation algorithm [28].
We do not consider these additions as contributions per se, but describing them here is required to understand the experimental contribution of this paper.\\
Indeed, thanks to these tools we have been able to design a complete pipeline for synthesizing multi-contact motions from scratch.
The novelty of the present paper lies in the rigorous experimental setup we propose to demonstrate its interest.
As opposed to our previous paper, we demonstrate the algorithms with actual robotic models. These models are much more constrained kinematically than the virtual avatars used before,
thus making the planning problem even harder. Through an extensive empirical analysis, we demonstrate that the method remains interesting in this context, qualitatively and efficiently.}\done
